---
title: 'Lab 11: Comparing Binomial Proportions'
author: "Jishan Luo"
date: "April 1, 2019"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A good reference book
The following problem is taken from page 105 `Bayesian Ideas and Data Analysis: An Introduction for Scientists and Statisticians` 



# The story

534 medflies that lived at least 34 days. These are cross classified by two factors:

    A) whether the fly produced at least 1,000 eggs in the first 30 days of life and
    B) whether the fly lived ata least 44 days
    
```{r}
mat = matrix(c(54,80,224,176), nr = 2,nc=2, byrow = TRUE)
colnames(mat)= c("Long-lived","short-lived")
rownames(mat)= c("High", "Low")
mat
```
These are multinomial data but if we view the Longetivity given reproductive factor we shall see these as two independent binomials

$$y_1|\theta \sim Bin(n_1,\theta_1) \perp y_2|\theta_2 \sim Bin(n_2, \theta_2)$$
    
The data can then be obtained from the above table:

> 
    $n_1=134,n_2=400, y_1=54, y_2 = 224$
    
    
    
# Task 1

Find the conjugate solution -- that is find the *analytical* posteriors for $\theta_1|y_1$ and $\theta_2|y_2$ using $\theta_i\sim Beta(a_i,b_i)$ where $i\in \{1,2\}$

$$p(\theta_1|y_1)=Beta(a_1+y_1,b_1+n_1-y_1)$$
$$p(\theta_2|y_2)=Beta(a_2+y_2,b_2+n_2-y_2)$$


# Task 2
Make plots of the two posteriors in R using

* $a_i = 1, b_i =1$
* $a_i = 4, b_i =1$
* $a_i = 1, b_i =4$
* $a_i = 3, b_i =3$

```{r }
#layout(matrix(1:8,nc=2,nr=4))
a=c(1,4,1,3)
b=c(1,1,4,3)
n=c(134,400)
y=c(54,224)
for (i in 1:2)
{
  for(j in 1:4)
  {
    curve(dbeta(x,a[j]+y[i],b[j]+n[i]-y[i]),xlim=c(0,1),lwd=2,col=c("Red","Green")[i],ylab="density",main="Jishan")
    text(0.3,6,paste("a=",a[j],"b=",b[j]))
  }
}
```



# Task 3

Create a jags model for the above problem by correcting and/or completing the given code. Note that there are some useful logical nodes made. Comment on them.

```{r model, eval=FALSE}
model{
  for(i in 1:2){
  y[i] ~ dbin(theta[i], n[i])

  theta[i] ~ dbeta(1,1)
  odds[i] <- theta[i]/(1-theta[i]) 
  }
  
  RD <- theta[2]-theta[1] # Risk difference
  RR <-theta[2]/theta[1] #Relative risk
  test<- step(RD) # What does this test? It helps us to find the mean of the test. The mean of the test is the probability of RD is positive. 
}
```



# Task 4

Now make the complete jags code. Please include some logical constant assignments so that you can assign new priors easily

Example:

```{r logicalnodes, eval=FALSE}
a[i]<-1
b[i]<-1
```


Make sure you monitor `theta`, `RR`, `OR` and `test`.
```{r jagscript}
library(rjags)

a=1 #prior 
b=1

dataList=list(n=n,y=y,a=a,b=b) #this can change the hyper-parameters

#Define the model:
modelString = "
model{
for(i in 1:2){
  y[i] ~ dbin(theta[i], n[i])

  theta[i] ~ dbeta(a,b) #prior
  odds[i] <- theta[i]/(1-theta[i]) 
  }
  
  RD <- theta[2]-theta[1] # Risk difference
  RR <-theta[2]/theta[1] #Relative risk
  test<- step(RD)

}

" # close quote for modelString
writeLines( modelString , con="TEMPmodel.txt" )

#  initsList = list( theta=thetaInit )

initsList = list(theta=c(0.4,0.5))
# Run the chains:
jagsModel = jags.model( file="TEMPmodel.txt" , data=dataList , inits=initsList , 
                        n.chains=3 , n.adapt=500 )
list.samplers(jagsModel)

update( jagsModel , n.iter=500 )
codaSamples = coda.samples( jagsModel , variable.names=c("theta", "RR","RD","test"),n.iter=33340 )
save( codaSamples , file=paste0("lab11","Mcmc.Rdata") )

summary(codaSamples)

library(ggmcmc)
s = ggs(codaSamples)
d = ggs_density(s)

print(d)

cr = ggs_crosscorrelation(s)
print(cr)
```



# Task 5

Please give point and interval estimates for the above nodes. Interpret them.
```{r}
summary(codaSamples)
```
>Point estimates:

The mean value of `RD` is   0.1553;

The mean value of `RR` is   1.3990;

The mean value of `test` is    0.9993;

The mean value of `theta[1]`  is   0.4044;

The mean value of `theta[2]` is   0.5597.



>Interval estimates:

Given a 95% Baysian credible interval for `RD`, we got the interval value is ( 0.05968,0.2489). We are 95% confidet that the true value of `RD` is contained in that interval;

Given a 95% Baysian credible interval for `RR`, we got the interval value is (1.12559,1.7506). We are 95% confidet that the true value of `RR` is contained in that interval;

Given a 95% Baysian credible interval for `test`, we got the interval value is (1.00000,1.0000). We are 95% confidet that the true value of `test` is contained in that interval;

Given a 95% Baysian credible interval for `theta[1]`, we got the interval value is (0.32428,0.4871). We are 95% confidet that the true value of `theta[1]` is contained in that interval;

Given a 95% Baysian credible interval for `theta[2]`, we got the interval value is (0.51071,0.6075). We are 95% confidet that the true value of `theta[2]` is contained in that interval.


# Task 6

Make a function `mybibin()` that will solve the same type of problem where the data inputs will be a vector of y's and a vector of n's. The priors will be beta's and will therefore need a vector of a's and b's. 
The command line output should be a list of point and interval estimates for a given $\alpha$.
The function should produce analytical posterior, prior and likelihood plots and MCMC posterior plots of the theta's and all other stochastic nodes.
Run your function with the same data as given and $a = c(1,2), b= c(1,2)\; and\; \alpha = 0.2$

How will you change the model code? 
There are many ways to do this -- one way is to break the model "text" into parts and rejoin them after changing values. You can use `paste()`

```{r}
mybibin=function(y=c(54,224),n=c(134,400),alpha=0.05,a=c(1,2),b=c(1,2)){
  
library(rjags)
library(ggmcmc)
dataList=list(n=n,y=y,a=a,b=b) #this can change the hyper-parameters

#Define the model:
modelString = "
model{
for(i in 1:2){
  y[i] ~ dbin(theta[i], n[i])

  theta[i] ~ dbeta(a[i],b[i]) #prior
  odds[i] <- theta[i]/(1-theta[i]) 
  }
  
  RD <- theta[2]-theta[1] # Risk difference
  RR <-theta[2]/theta[1] #Relative risk
  test<- step(RD)

}

" # close quote for modelString
writeLines( modelString , con="TEMPmodel.txt" )

#  initsList = list( theta=thetaInit )

initsList = list(theta=c(0.4,0.5))
# Run the chains:
jagsModel = jags.model( file="TEMPmodel.txt" , data=dataList , inits=initsList , 
                        n.chains=3 , n.adapt=500 )
list.samplers(jagsModel)

update( jagsModel , n.iter=500 )
codaSamples = coda.samples( jagsModel , variable.names=c("theta", "RR","RD","test"),n.iter=33340 )
save( codaSamples , file=paste0("lab11","Mcmc.Rdata") )

summary(codaSamples)

library(ggmcmc)
s = ggs(codaSamples)
d = ggs_density(s)

print(d)

cr = ggs_crosscorrelation(s)
print(cr)

t=ggs_traceplot(s)
print(t)

list(summary=summary(codaSamples,c(alpha/2,1-alpha/2)),data=dataList,density=d)

}


mcout2=mybibin(alpha=0.20,a=c(1,2),b=c(1,2))
mcout2$summary
mcout2$density
mcout2$data
```

