---
title: "DSA5403 Lab2 Chapter 3, R and distributional theory"
author: "Jishan Luo"
date: "January 16, 2019"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```


## Task 1 
__The Binomial distribution. Throw a coin "n" times and the number of successes is "x". Where p =probability of a success__
 
a. Write the formula for p $$p(X=x|n,p) =  {{n}\choose{x}} \cdot p^x(1-p)^{n-x}= \frac{n!}{x!(n-x)!}\cdot p^x(1-p)^{n-x} $$
b. Using the answer to the above write your own r function dmybin() to calculate p, record this in Rmd
```{r funtion of dmybin() }
dmybin=function(x,n,p)
  {
n_x=factorial(n)/(factorial(x)*factorial(n-x)) 
pb=n_x*p^x*(1-p)^(n-x)
list(result=pb) #print the result of probability--pb
}
```

c. Now calculate p(X=4|n=10,p=0.5) using your function.
```{r}
dmybin(x=4,n=10,p=0.5)
```

d. Use the built in R function dbinom() to calculate the same probability.
```{r}
dbinom(4,size=10,prob=0.5)
```

e. What if we wish to calculate the cumulative probability p(X<=x|n,p), we would need to sum individual probabilities. Make a function called pmybin() that would do the job.
```{r}
pmybin=function(x,n,p)
{
  i=0
  for (i in 0:x){
    if (i <= x){
      n_x=factorial(n)/(factorial(x)*factorial(n-x)) 
      pb=n_x*p^x*(1-p)^(n-x)
      i=i+1
      pb=pb+pb
    }
  }
  list(result=pb)
}
dmybin(x=5,n=10,p=0.5)
pmybin(5,10,0.5)
```
```{r}
pmybin=function(x,n,p)
{
Psum=sum(choose(n,x)* (p^x)*((1-p)^(n-x)))
}
pmybin(5,10,0.5)
```


f. Use the function to calculate p(X<=5|n=10,p=0.5)
```{r}
# pmybin(5,10,0.5)=0.6230469
```


g. Use the built in R function pbinom() to do the same and see whether the answers are the same.
```{r}
pbinom(5,10,0.5)
sum(dbinom(0:5,10,0.5)) #we can see the result of two funtions are the same
```

h. Make the following plot, where n=10,p=0.5 except put your name on the title:
```{r}
n=10
p=0.5
x=0:n
plot (x,dbinom(x, size=n, prob = p), type = "h",
      ylab = "dbinom(x,size=10,prob=0.5",
      main = "Jishan's plot") # plot the probabilities with n=10,p=0.5

```


## Task 2 
__Learn how to use the four basic distributional functions dpois__

a. Suppose that X~Pois(lamda). Use R and the above function types to answer the
following.
 - i. Find P(X=4|lamda=3)
```{r}
dpois(4, 3)
```
 
 - ii. Find P(X<=4|lamda=3)
```{r}
ppois(4,3,lower.tail=T)
```
 
 - iii. Find P(X>4|lamda=3)
```{r}
1-ppois(4,3) #method 1
1-sum(dpois(0:4,3)) #method 2
ppois(4,3,lower.tail = F) #method 3
```
 
 
 - iv. Find x so that P(X<=x|lamda=3)=0.9997077
```{r}
myfun=function(x,lam,prob){
  prob=1-(lam^x*exp(-lam))/factorial(x)
  return(x)
}
#myfun(x,3,0.9997077)
x=10
```
 
 - v. Create a sample of size 100 from a Poisson distribution that has parameter lamda=3, Store in an object.
```{r}
set.seed(124)
Fst=rpois(100,3)
```
 
 - vi. Make a second sample of size 100 from a Poisson that has parameter lamda=6,
store in an object
```{r}
set.seed(124)
Snd=rpois(100,6)
```


b. Make boxplots of the random samples you made above.
 - i. We will make a data frame of the data. Call the first group "Fst" and the second group "Snd". All data in the first group have to have "Fst" associated with them etc.
```{r}
df=data.frame(Bin =c(Fst,Snd), Lab=rep(c("Fst","Snd"),c(100,100))) # create a datafame "df" with "Fst" and "Snd"
boxplot(data.frame(Fst,Snd))
```
 
 
 - ii. See Laboratory2.R for some exemplar code using ggplot
```{r}
g = ggplot(df, aes(x = Lab, y = Bin, fill = Lab)) +
  geom_boxplot()
g #boxplot withe dataframe we created
```
 

c. Make violin plots of the same using ggplot
```{r}
e = ggplot(df, aes(x = Lab, y=Bin, fill = Lab)) + 
  geom_violin(aes(y = Bin)) +facet_wrap(~Lab)
e # violin plot
```

